CRUSE to Safe Cycling in Ireland \\\large{An Open Source Tool to Support Active Travel}

### 2023-10-10 paper.qmd
Final tweaks

In Ireland, the Road Safety Authority (RSA) has set the target of halving the number of road traffic deaths and serious injuries by 2030 [@national2021].
Doing so while simultaneously enabling rapid uptake of active modes will require key active travel routes to be identified and improved.

A proactive approach to the development of cycling policy and infrastructure is taking place in Ireland.
The main frameworks underpinning these efforts are the Climate Action Plan [@climate2022a], the National Development Plan, and the National Roads 2040 strategy [@national2023].
<!-- Strategic alignment is also outlined by the Phoenix Park Transport and Mobility Options Study, which highlights public support for active travel interventions: while only 3% of trips in Ireland are made by cycling according to the latest Census data from 2016,^[ --> These policy frameworks are backed-up by public support for active travel interventions: while only 3% of trips in Ireland are made by cycling according to the latest Census data from 2016 "83% of respondents supported enhanced walking and cycling facilities" [@sustaina2019].

@@

The aim of the paper was to describe the design, features and potential use of the CRUSE Tool.
As outlined in the previous section, the CRUSE Tool provides evidence on current and potential future cycling levels across Ireland down to the street level, with potential assigned to 'fastest', 'balanced' and 'quietest' route networks.
This provision of multiple scenarios of behaviour change *and* multiple scenarios of investment, for example in cycling infrastructure next to major roads vs quiet residential streets, is a key feature of the CRUSE Tool.

@@

While the CRUSE Tool is not the first national and publicly available tool for cycling planning, it has some key features that make it relevant for other countries, and other national road authorities tasked with making their transport systems safe:

@@

@@

The CRUSE Tool is an open access web application for strategic cycle network planning and prioritization of road safety interventions across Ireland.
Building on previous work, it provides a nationally consistent evidence base, providing valuable insights to planners and other stakeholders in the transport planning process, at national and local levels.
Because the results are available at the route segment level, the tool can be used to identify 'weak links' in the cycle network, and to prioritize investment in cycling infrastructure.
Furthermore, the provision of the evidence in a free and publicly available website, hosted at [cruse.bike](https://cruse.bike), means that it can be used by anyone, encouraging wider participation and more evidence-based debate about transport planning.

>>>>

In Ireland, the Road Safety Authority (RSA) has set the target of halving the number of road traffic deaths and serious injuries by 2030 [@national2021].
Doing so while simultaneously enabling rapid uptake of active modes will require key active travel routes to be identified and improved.
A proactive approach to the development of cycling policy and infrastructure is taking place in Ireland.
The main frameworks underpinning these efforts are the Climate Action Plan [@climate2022a], the National Development Plan, and the National Roads 2040 strategy [@national2023].
<!-- Strategic alignment is also outlined by the Phoenix Park Transport and Mobility Options Study, which highlights public support for active travel interventions: while only 3% of trips in Ireland are made by cycling according to the latest Census data from 2016,^[ --> These policy frameworks are backed-up by public support for active travel interventions: while only 3% of trips in Ireland are made by cycling according to the latest Census data from 2016 "83% of respondents supported enhanced walking and cycling facilities" [@sustaina2019].

@@

The aim of the paper was to describe the design, features and potential use of the CRUSE Tool.
As outlined in the previous section, the CRUSE Tool provides evidence on current and potential future cycling levels across Ireland down to the street level, with potential assigned to 'fastest', 'balanced' and 'quietest' route networks.
This provision of multiple scenarios of behavior change *and* multiple scenarios of investment, for example in cycling infrastructure next to major roads vs quiet residential streets, is a key feature of the CRUSE Tool.

@@

While the CRUSE Tool is not the first national and publicly available tool for cycling planning, it has some key features that make it relevant for other countries, regions and road authorities tasked with making their transport systems safe and sustainable:

@@

As the benefits of strategic planning for cycling become more apparent [@scappini2022] we expect the demand for tools such as CRUSE to grow.

@@

The CRUSE Tool is an open access web application for strategic cycle network planning and prioritization of road safety interventions across Ireland.
Building on previous work, it provides a nationally consistent evidence base, providing valuable insights to planners and other stakeholders in the transport planning process, at national and local levels.
Because the results are available at the route segment level, the tool can be used to identify 'weak links' in the cycle network [@vybornova2022], and to prioritize investment in cycling infrastructure to maximize health, equality, and other benefits [@mahfouz; @woodcock2021].
Furthermore, the provision of the evidence in a free and publicly available website, hosted at [cruse.bike](https://cruse.bike), means that it can be used by anyone, encouraging wider participation and more evidence-based debate about transport planning.

### 2023-10-10 paper.tex
Final tweaks

For investment in cycling to deliver on the huge potential benefits,
cycling must become safer. With growth in the E-bike market, CRUSE will
help inform inter-urban and rural networks to support the transfer of
trips to sustainable modes for longer journeys. Provision of open access
data in a publicly available web tool will address the challenges faced
by counties across Ireland in meeting their sustainability commitments.
The experience of developing a deploying CRUSE are relevant to countries
with sustainable transport policies worldwide.
\end{abstract}

\begin{keyword}
    Cycling \sep Open-source \sep Road Safety \sep 
    Active Travel
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, enhanced, frame hidden, boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners]}{\end{tcolorbox}}\fi

@@

Uptake of cycling should be seen in this broader context of transport
decarbonisation \citep{brand2020} and sustainable mobility
\citep{burns2013}. Although cycling uptake appears on the surface to
only relate to the `shift' part of the ASI framework, closer
consideration of the knock-on impacts of cycling uptake shows that it
can also help avoid unnecessary trips \citep{nello-deakin2020}.
Furthermore, highly efficient e-bikes --- which are seeing rapid uptake
in Ireland and other countries --- outperform electric cars, which are
too heavy and expensive. Over-reliance on electric cars could slow the
transition away from car dependency and inadvertently enable ``high
travel lock-in'' \citep{anable2019}. At the European level, the European
Union has a target of reducing GHG emissions by 55\% by 2030, compared
to 1990 levels, and to achieve `net-zero' by 2050 \citep{rosenow2022}.

Another deadly consequence of inefficient transport systems is tackled
in the Road Infrastructure Safety Management (RISM) directive
(2008/96/EC), which requires member states to implement a road safety
management system (RSMS) for all public roads. Specifically, ``Member
States shall ensure that the ranking of high accident concentration
sections and the network safety ranking are carried out''
\citep{directiv2008}. Given that `safety' in this context is usefully
quantified as the number of people killed and seriously injured (KSI)
per distance traveled, the directive requires estimation of distance
traveled by mode, down to the road link level. For active modes, about
which there is a paucity of data compared with motorized modes, this is
a major challenge. Better data to inform road safety policies and
interventions is a motivation for the CRUSE Tool outlined in this paper,
and modeling active travel more broadly.

In Ireland, the Road Safety Authority (RSA) has set the target of
halving the number of road traffic deaths and serious injuries by 2030
\citep{national2021}. Doing so while simultaneously enabling rapid
uptake of active modes will require key active travel routes to be
identified and improved.

A proactive approach to the development of cycling policy and
infrastructure is taking place in Ireland. The main frameworks
underpinning these efforts are the Climate Action Plan
\citep{climate2022a}, the National Development Plan, and the National
Roads 2040 strategy \citep{national2023}. These policy frameworks are
backed-up by public support for active travel interventions: while only
3\% of trips in Ireland are made by cycling according to the latest
Census data from 2016 ``83\% of respondents supported enhanced walking
and cycling facilities'' \citep{sustaina2019}.

At the regional level, the recently published Greater Dublin Area
Transport Strategy \citep{greater} reinforces these findings: ``nearly a
quarter of adults cycle at least once a week in the Dublin Metropolitan
Area'' with cycling in the Dublin area taking up to 60,000 cars off the
road today. Extrapolating this on a per population basis across Ireland,
with around 40\% of the population living in Dublin, suggests that
around 150,000 cars could be removed nationwide just by achieving Dublin
levels of cycling in all counties (notwithstanding existing cycling
trips and differences in trip distances). 7 in 10 trips in Ireland are
by car \citep{national} . Cycling has the potential to replace a large
proportion of these trips: ``A high priority must also be given to
cyclists, because trips by this mode have the potential to replace trips
by private car, most specifically for short to medium distance trips,
but increasingly for longer trips as e-bikes extend the range of this
mode'' \citep{greater}.

@@

The aim of the paper was to describe the design, features and potential
use of the CRUSE Tool. As outlined in the previous section, the CRUSE
Tool provides evidence on current and potential future cycling levels
across Ireland down to the street level, with potential assigned to
`fastest', `balanced' and `quietest' route networks. This provision of
multiple scenarios of behaviour change \emph{and} multiple scenarios of
investment, for example in cycling infrastructure next to major roads vs
quiet residential streets, is a key feature of the CRUSE Tool.

@@

@@

The CRUSE Tool is an open access web application for strategic cycle
network planning and prioritization of road safety interventions across
Ireland. Building on previous work, it provides a nationally consistent
evidence base, providing valuable insights to planners and other
stakeholders in the transport planning process, at national and local
levels. Because the results are available at the route segment level,
the tool can be used to identify `weak links' in the cycle network, and
to prioritize investment in cycling infrastructure. Furthermore, the
provision of the evidence in a free and publicly available website,
hosted at \href{https://cruse.bike}{cruse.bike}, means that it can be
used by anyone, encouraging wider participation and more evidence-based
debate about transport planning.

>>>>

For investment in cycling to deliver on the huge potential benefits,
cycling must become safer. With growth in the ebike market, CRUSE will
help inform inter-urban and rural networks to support the transfer of
trips to sustainable modes for longer journeys. Provision of open access
data in a publicly available web tool will address the challenges faced
by counties across Ireland in meeting their sustainability commitments.
The experience of developing a deploying CRUSE are relevant to countries
with sustainable transport policies worldwide.
\end{abstract}

\begin{keyword}
    Cycling \sep Open-source \sep Road Safety \sep 
    Active Travel
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, frame hidden, sharp corners, breakable, enhanced]}{\end{tcolorbox}}\fi

@@

Building cycle networks represents a relatively `quick win' within the
context of decarbonization \citep{brand2020} and sustainable mobility
\citep{burns2013}. Although cycling uptake appears on the surface to
only relate to the `shift' part of the ASI framework, closer
consideration of the knock-on impacts of cycling uptake shows that it
can also help avoid unnecessary trips \citep{nello-deakin2020}.
Furthermore, highly efficient ebikes --- which are seeing rapid uptake
in Ireland and other countries --- outperform electric cars, which are
too heavy and expensive. Over-reliance on electric cars could slow the
transition away from car dependency and inadvertently enable ``high
travel lock-in'' \citep{anable2019}. At the European level, the European
Union has a target of reducing GHG emissions by 55\% by 2030, compared
to 1990 levels, and to achieve `net-zero' by 2050 \citep{rosenow2022}.
Climate change mitigation is a major motivation for cycle network plans
\citep{scappini2022}.

Another motivation for cycle network planning is the Road Infrastructure
Safety Management (RISM) directive (2008/96/EC), which requires member
states to implement a road safety management system (RSMS) for all
public roads. Specifically, ``Member States shall ensure that the
ranking of high accident concentration sections and the network safety
ranking are carried out'' \citep{directiv2008}. Given that `safety' in
this context is usefully quantified as the number of people killed and
seriously injured (KSI) per distance traveled, the directive requires
estimation of distance traveled by mode, down to the road link level.
For active modes, about which there is a paucity of data compared with
motorized modes, this is a major challenge. Better data to inform road
safety policies and interventions is a motivation for the CRUSE Tool
outlined in this paper, and modeling active travel more broadly.

In Ireland, the Road Safety Authority (RSA) has set the target of
halving the number of road traffic deaths and serious injuries by 2030
\citep{national2021}. Doing so while simultaneously enabling rapid
uptake of active modes will require key active travel routes to be
identified and improved. A proactive approach to the development of
cycling policy and infrastructure is taking place in Ireland. The main
frameworks underpinning these efforts are the Climate Action Plan
\citep{climate2022a}, the National Development Plan, and the National
Roads 2040 strategy \citep{national2023}. These policy frameworks are
backed-up by public support for active travel interventions: while only
3\% of trips in Ireland are made by cycling according to the latest
Census data from 2016 ``83\% of respondents supported enhanced walking
and cycling facilities'' \citep{sustaina2019}.

At the regional level, the recently published Greater Dublin Area
Transport Strategy \citep{greater} reinforces these findings: ``nearly a
quarter of adults cycle at least once a week in the Dublin Metropolitan
Area'' with cycling in the Dublin area taking up to 60,000 cars off the
road today. Extrapolating this on a per population basis across Ireland,
with around 40\% of the population living in Dublin, suggests that
around 150,000 cars could be removed nationwide just by achieving Dublin
levels of cycling in all counties (notwithstanding existing cycling
trips and differences in trip distances). 7 in 10 trips in Ireland are
by car \citep{national} . Cycling has the potential to replace a large
proportion of these trips: ``A high priority must also be given to
cyclists, because trips by this mode have the potential to replace trips
by private car, most specifically for short to medium distance trips,
but increasingly for longer trips as ebikes extend the range of this
mode'' \citep{greater}.

@@

The aim of the paper was to describe the design, features and potential
use of the CRUSE Tool. As outlined in the previous section, the CRUSE
Tool provides evidence on current and potential future cycling levels
across Ireland down to the street level, with potential assigned to
`fastest', `balanced' and `quietest' route networks. This provision of
multiple scenarios of behavior change \emph{and} multiple scenarios of
investment, for example in cycling infrastructure next to major roads vs
quiet residential streets, is a key feature of the CRUSE Tool.

@@

As the benefits of strategic planning for cycling become more apparent
\citep{scappini2022} we expect the demand for tools such as CRUSE to
grow.

@@

The CRUSE Tool is an open access web application for strategic cycle
network planning and prioritization of road safety interventions across
Ireland. Building on previous work, it provides a nationally consistent
evidence base, providing valuable insights to planners and other
stakeholders in the transport planning process, at national and local
levels. Because the results are available at the route segment level,
the tool can be used to identify `weak links' in the cycle network
\citep{vybornova2022}, and to prioritize investment in cycling
infrastructure to maximize health, equality, and other benefits
\citep{mahfouz, woodcock2021}. Furthermore, the provision of the
evidence in a free and publicly available website, hosted at
\href{https://cruse.bike}{cruse.bike}, means that it can be used by
anyone, encouraging wider participation and more evidence-based debate
about transport planning.
